### 2022-10-22 dissertation/chapters/02-artificial-neural-networks.tex
🚧 WIP - Save point (compiles)

\noindent
The human brain is one of the most complicated biological organs in nature. It gives us the ability to think and learn. The field of \acs{ML} has taken a lot of inspiration from the biological brain which lead to the development of \acp{ANN} \cite{ref:rosenblatt:1958}. \acp{ANN} are used throughout this dissertation as the \textit{model} to be trained using \acp{HH}.

@@

\noindent
The biological neural systems are made up of microscopic nerve cells called neurons \cite{ref:jain:1996}. Figure~\ref{fig:biological_neuron} illustrates a single \acs{BN}. The main components of the \acs{BN} include the cell body, \index{dendrites}\textit{dendrites} and the \index{axon}\textit{axon}. \Acp{NN} are formed through the connections between the \index{axon}axons and \index{dendrites}dendrites of various neurons. This is known as \index{synaptogenesis}\textit{synaptogenesis} \cite{ref:huttenlocher:1997}. Such a connection is referred to as a \index{synapse}\textit{synapse}. Communication takes place, through the synapse, by electro-chemical pulse and is often referred to as an \textit{activation} or \textit{action potential}.  Communication signals propagate from the \index{dendrites}dendrites, through the cell body to the axon of a neuron, provoking a signal in the post-synaptic neuron \cite{ref:engelbrecht:2007}. The greater the connection between two neurons, the stronger the communication.  Kennedy \cite{ref:kennedy:2016} defines stronger \index{synapse}synapses as ones that contribute more depolarisation to the neural membrane upon activation than weaker ones. Stronger \index{synapse}synapses have a higher probability of generating an action potential in the post-synaptic neuron. During activation, the pre-synaptic neuron release neurotransmitters that bind to the post-synaptic neuron \cite{ref:khanacademy:synapse}. The frequent release of these molecules cause the \index{synapse}synapse to grow. Connections that grow over time yield stronger signals (learning), while connections that are weak propagate low intensity signals and vanish over time (forgetting). The ability of synapses to strengthen and weaken over time is known as \index{synaptic plasticity}\textit{synaptic plasticity} \cite{ref:huttenlocher:1997}.

@@

The \acs{AN} implements a nonlinear mapping from $\mathbb{R}^{I}$ to
$\mathbb{R}^{T}$, usually in the ranges $[0,1]$ or $[-1,1]$, depending on the
activation function used \cite{ref:engelbrecht:2007} and is given
in Equation~\eqref{eq:an_function_mapping} below.

@@

Input signals are $I$-dimensional vectors of numerical values that are obtained either through some environment sensor or from other \acp{AN}. Input signals are often referred to as \index{features}\textit{features} or \index{independent variables}\textit{independent variables} \cite{ref:francis:2001}. Throughout this dissertation, a single input vector is referred to as a \textit{pattern}. Input data must first be pre-processed before it is presented to the \acs{AN}. Input pre-processing techniques are presented in the following sections.

@@

Input pre-processing improves the training process and contributes to the success of the practical application of the \acs{ANN} \cite{ref:kuzniar:2017}. The primary purpose of input pre-processing is to modify the input data so that it can better match predicted output data. There are many techniques to consider when pre-processing input data. These techniques include methods to encode data to certain formats, to scale to the correct ranges and to deal with incomplete, invalid or irrelevant data. For the purposes of this dissertation, \textit{data encoding} and \textit{normalisation} techniques are considered.

@@

For qualitative data, the class labels have to be converted from textual to numerical representations. In general, class labels are encoded as numerical vectors \cite{ref:srinidhi:2018, ref:brownlee:2017:one-hot}. One such encoding technique is referred to as \index{one-hot encoding}\textit{one-hot encoding}. Harris and Harris \cite{ref:harris:2010} describe \index{one-hot encoding}one-hot encoding as a group of vector elements where the classes are represented with a single high element (represented by $1$) and all the others low (represented by $0$). Each element $y_k$, where $k$ is the $k$-th element of the encoded vector $\vec{y}$, uniquely refers to a class. An activation, represented by a value of $1$ at a given index thus represents a class associated with that index.

@@

Numerical data must be normalised and scaled appropriately for the \acs{AN}. In general, input data is normalised to a range that is appropriate for the \index{activation function}activation function being used by the \acs{AN}. Normalisation and scaling yields input data that is comparable to the output produced by the \acs{AN}. For the purposes of this disseration, the \index{min-max scaler}min-max scaler \cite{ref:al:2006} and the \index{standard score scaler}standard score scaler \cite{ref:jain:2005} are considered.

@@

The connection strength that synapses in the \acs{BN} have is modeled in the \acs{AN} as weight vectors of numerical values associated with each dimension of the input. Weights can either dampen or strengthen the input by some negative or positive numerical value. Changes in the weight associated with a feature changes the influence that that particular feature has on the predicted output. Finding the correct value for each weight, such that the \acs{AN} yields optimal output, is an optimisation problem \cite{ref:thierens:1993}. Research has shown that weight initialisation plays an important role in the efficient training of \acp{ANN} \cite{ref:thimm:1995}.

@@

Weight initialisation is the process by which candidate solutions (represented by the weights of the \acs{AN}) to the problem are ``placed'' in the search-space. Weight initialisation influences the speed of convergence, the probability of convergence and the generalisation capabilities of \acp{ANN} \cite{ref:fernandez:2001}. Finding the optimal initialisation values for weights is non-trivial and can be seen as another optimisation problem \cite{ref:de:2016, ref:erdogmus:2003, ref:yam:2000}.

Weight initialisation is dependent on the \index{activation function}activation function used. If weights are initialised as values that are too small, the vanishing gradients problem can occur \cite{ref:hanin:2018}. If weights are initialised as values that are too big, output of the activation function would not be in the active range. This leads to unit saturation, and exploding gradients can occur \cite{ref:hanin:2018, ref:yadav:2018}.

Many different weight initialisation techniques have been developed \cite{ref:erdogmus:2003}. Some techniques have been developed to address the problems that are mentioned above \cite{ref:yadav:2018}. For the purposes of this dissertation, focus is put on \index{random uniform sampling}random uniform sampling, \index{Glorot uniform sampling}\textit{Glorot uniform (Xavier)} sampling and \index{Glorot normal sampling}\textit{Glorot normal} sampling \cite{ref:glorot:2010}. Brief discussions on each of these weight initialisation techniques are presented as follows.

@@

\index{random uniform sampling}Random uniform sampling initialises weights uniformly in the range $[\omega_{min}, \omega_{max}]$, written as $w_{i} \sim \textit{U} (\omega_{min}, \omega_{max})$, where the $\omega_{min}$ and $\omega_{max}$ are respectfully the lower and upper bounds of the uniform distribution. Suggested parameter values are $(-1, 1)$ or $(-0.5, 0.5)$ \cite{ref:nguyen:1990}.

@@

\index{Glorot normal sampling}Glorot normal sampling initialises weights by sampling from a truncated normal distribution centred on a mean of $0$ and with $\sigma = \sqrt{\frac{2}{I + K}}$, where $\sigma$ is the standard deviation of the distribution. $I$ and $K$ are respectfully the number of input and output units in the weight vector. \index{Glorot normal sampling}Glorot normal sampling has been shown to decrease training time \cite{ref:glorot:2010}.

@@

A common way by which the \index{net input signal}net input signal is calculated is by means of \acp{SU} \cite{ref:engelbrecht:2007}, which compute the \index{net input signal}net input signal as the weighted sum of all input signals as shown in Equation~\eqref{eq:summation_units} below.

@@

A bias/threshold term $\theta$ is introduced to help translate the output of the \index{activation function}activation function \cite{ref:benitez:1997}. The value of $\theta$ can be learned during the training process along with the weights of the \acs{ANN}. In order to simplify equations, the input and weight vectors are augmented such that the input and hidden layers have an additional neuron/unit, called the \textit{bias unit} \cite{ref:engelbrecht:2007}. The net input signal can then be rewritten to consider the bias unit, leading to an augmented net input signal.

@@

An \index{activation function}activation function is used to model the action potential of the \acs{AN} \cite{ref:ziv:1994, ref:hodgkin:1952}. The \index{activation function}activation function takes as a parameter, the augmented \index{net input signal}net input signal. When the output produced by the activation function surpasses some threshold value $\tau$, we consider that neuron to have ``fired'' an output signal. \index{activation function}Activation functions thus model \textit{phase shift}. In the context of classification problems, \index{activation function}activation functions form decision boundaries between classes. In the context of regression problems, \index{activation function}activation functions try to approximate some function that maps the input data to some target data.

@@

\noindent
Many different \index{activation function}activation functions have been developed \cite{ref:karlik:2011}. For the purposes of this dissertation, focus is put on the \acf{ReLU} \cite{ref:jarrett:2009, ref:nair:2010}, the \acf{LReLU} \cite{ref:maas:2013}, the \index{sigmoid}sigmoid \cite{ref:lecun:1988} and the \index{hyperbolic tangent}hyperbolic tangent \cite{ref:lin:2008} \index{activation function}activation functions.

@@

\noindent
An advantage of \acs{ReLU} is that it is not susceptible to the \index{vanishing gradients problem}vanishing gradients problem \cite{ref:xu:2015, ref:maksutov:2018} which occurs when gradients in the first layers of a multi-layer \acs{ANN} approach zero, and have no effect on the training process.

@@

The \acf{LReLU} \index{activation function}activation function is a variant of the \acs{ReLU} \index{activation function}activation function that avoids zero gradients in the negative part of its argument by introducing a scaling parameter, $\alpha > 0$ \cite{ref:xu:2015}. Similar to \acs{ReLU}, \acs{LReLU} is not susceptible to the \index{vanishing gradients problem}vanishing gradients problem. However, \acs{LReLU} does not suffer from the \index{dying ReLU problem}dying \acs{ReLU} problem \cite{ref:agarap:2018}, which occurs when the neurons that use \acs{ReLU} \index{activation function}activation functions become \textit{inactive} and only output $0$ due to a negative net input signal. The \acs{LReLU} \index{activation function}activation function is defined in Equation~\eqref{eq:leaky_relu} below.

@@

\noindent
By introducing a parameter $\alpha > 0$, negative \index{net input signal}net input signals will still yield non-zero activations, resulting in non-zero gradients and avoiding gradient saturation. Non-zero gradients are required by some heuristics such as \acs{SGD} in order to be able to effectively train \acp{ANN} \cite{ref:hanin:2018}. An illustration of \acs{LReLU} with various values for $\alpha$ and $\theta = 0$ is given in Figure~\ref{fig:anns:activation_functions:leaky_relu}.

@@

The \index{sigmoid}sigmoid activation is the continuous differentiable approximation of the step function, which was used in the original perceptron model developed by \citeauthor{ref:rosenblatt:1957} \cite{ref:rosenblatt:1957}, and yields output in the range $(0, 1)$. The \index{sigmoid}sigmoid \index{activation function}activation function is defined in Equation~\eqref{eq:sigmoid} below.

@@

The \index{softmax}softmax function, also known as the \index{softargmax}\textit{softargmax} \cite[p.~184]{ref:goodfellow:2016} or \index{normalised exponential function}\textit{normalised exponential function} \cite{ref:bishop:2006} is a generalisation of the logistic function that converts a $K$-dimensional output vector $\vec{y}$ into a $K$-dimensional output vector $\vec{y^{'}}$ where each element $y^{'}_k$ is in the range $(0,1)$ and all elements sum up to $1$ as is shown in Equation~\eqref{eq:sum_after_softmax} below.

@@

\acp{ANN} are inspired by the biological brain. Our brains model biological neural networks with neuron counts in the hundreds of billions and synapse counts in the trillions. To emulate the computational capability of the human brain is not yet possible on modern day hardware. \citeauthor{ref:sandberg:2008} \cite{ref:sandberg:2008} approximates a computational requirement of 256 000 terabytes/s to emulate the entire brain. Despite our shortage in hardware capabilities, \acp{ANN} have been successfully applied to a range of problem classes. \citeauthor{ref:engelbrecht:2007} \cite{ref:engelbrecht:2007} summarises some common problems that are solved using \acp{ANN} and can be listed as follows.

\begin{itemize}
    \item \textbf{Classification}: Predicting the class of an input vector \cite{ref:khan:2001}.

    \item \textbf{Pattern Matching}: Producing closely associated patterns based on an input vector \cite{ref:cannady:1998, ref:kumar:1994}.

    \item \textbf{Pattern Completion}: Completing the missing parts of an input vector \cite{ref:dayhoff:2001}.

    \item \textbf{Optimisation}: Producing optimal values of parameters in a optimisation problems \cite{ref:specht:1991}.

    \item \textbf{Data Mining}: Feature discovery in large datasets \cite{ref:singh:2009}.
\end{itemize}

@@

The input layer contains the input data to the \acs{ANN}. Since the input layer simply provides the input data, some literature do not consider the input layer as an actual part of the \acs{ANN} \cite{ref:engelbrecht:2007}.

@@

The hidden layer contains a collection of ``hidden'' \acp{AN}, also referred to as \textit{hidden units} or \textit{nodes}. Hidden units are used if the target data is not linearly separable \cite{ref:engelbrecht:2007}. It has been shown that \acp{ANN} that incorporate monotonically increasing differentiable \index{activation function}activation functions can approximate any continuous function with just one hidden layer, given that the hidden layer has enough hidden units \cite{ref:hornik:1989}.

@@

The topology of the \acs{ANN} refers to the way that layers of \acp{AN} are connected to each other. There are many different topologies \cite{ref:miikkulainen:2010}. For the purposes of this dissertation focus is put on a \index{fully connected topology}\textit{fully connected} topology where each \acs{AN} in one layer is connected to all \acp{AN} in the next, without any cycles \cite{ref:zell:1994}.

@@

\acp{FFNN} were the first and simplest type of \acp{ANN} developed \cite{ref:schmidhuber:2015} and implement input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement \index{fully connected topology}fully connected topologies. In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes. Depending on the optimisation algorithm used, error correction information can be ``backpropagated'' through the network. An illustration of a \acs{FFNN} is given in Figure~\ref{fig:ffnn} below.

@@

\index{supervised learning}Supervised learning is the process of training where the data that is presented to the \acs{FFNN} during training, includes the desired solution \cite{ref:geron:2017}.  The \acs{FFNN} learns the mapping function from the input to the target output \cite{ref:brownlee:2016}. The desired solutions are referred to as \textit{labels}. \index{supervised learning}Supervised learning can be used for both classification and regression problems.

The training data that is used during supervised learning, is split proportionally into a \textit{training} and \textit{validation} set. Data in the training set is used to train the \acs{FFNN} \cite{ref:james:2013} and update the weights, while the validation dataset is used for hyper-parameter tuning, where the parameters of the training process are altered, but not the weights of the \acs{FFNN}.

@@

\index{overfitting}Overfitting describes a scenario where the trained \acs{FFNN} performs well on training data, but does not generalise well to never before seen data from the test set \cite{ref:tetko:1995, ref:geron:2017}. Geron \cite{ref:geron:2017} describes \index{overfitting}overfitting as the case where \acs{FFNN} is too complex relative to the noisiness of the training data.

@@

Underfitting describes a scenario where the \acs{FFNN} is not able to effectively learn the underlying structure of the training data \cite{ref:tetko:1995, ref:geron:2017}. Geron \cite{ref:geron:2017} describe underfitting as the case where the \acs{FFNN} is too simple relative to the underlying structure of the training data.\\
\\
\noindent
There are two types of supervised learning algorithms based on when weights are updated \cite{ref:engelbrecht:2007}. These include stochastic training and batch
training.

@@

\index{stochastic training}Stochastic training, also known as \index{online learning}\textit{online learning}, is a \index{supervised learning}supervised learning variation whereby weights are adjusted after each training pattern presented. \index{stochastic training}Stochastic training benefits from shuffling data in the training dataset before presenting it to the \acs{FFNN}, in order to avoid overfitting or memorising the order in which patterns are presented \cite{ref:engelbrecht:2007}. It has been shown that shuffling the training data can speed up convergence
\cite{ref:bengio:2012}.

@@

Research suggests a trade-off between stochastic and batch training by making use of mini-batches \cite{ref:bengio:2012}. Mini-batch training is similar to batch training, however weights are updated after $\beta$ patterns have been presented, where $\beta$ is the mini-batch size.\\
\\
\noindent
Performance metrics are used during training on the training set and evaluation on the test set. There are many different performance measurements that can be used, however this dissertation focuses on performance measures related to loss. Loss is calculated using an error function. The following section presents the reader with more detail on the error functions that can be used during the training process.

>>>>

\noindent
The human brain is one of the most complicated biological organs in nature. It gives us the ability to think and learn. The field of \acs{ML} has taken a lot of inspiration from the biological brain which lead to the development of \acp{ANN}~\cite{ref:rosenblatt:1958}. \acp{ANN} are used throughout this dissertation as the \textit{model} to be trained using \acp{HH}.

@@

\noindent
The biological neural systems are made up of microscopic nerve cells called neurons~\cite{ref:jain:1996}. Figure~\ref{fig:biological_neuron} illustrates a single \acs{BN}. The main components of the \acs{BN} include the cell body, \index{dendrites}\textit{dendrites} and the \index{axon}\textit{axon}. \Acp{NN} are formed through the connections between the \index{axon}axons and \index{dendrites}dendrites of various neurons. This is known as \index{synaptogenesis}\textit{synaptogenesis}~\cite{ref:huttenlocher:1997}. Such a connection is referred to as a \index{synapse}\textit{synapse}. Communication takes place, through the synapse, by electro-chemical pulse and is often referred to as an \textit{activation} or \textit{action potential}.  Communication signals propagate from the \index{dendrites}dendrites, through the cell body to the axon of a neuron, provoking a signal in the post-synaptic neuron~\cite{ref:engelbrecht:2007}. The greater the connection between two neurons, the stronger the communication.  Kennedy~\cite{ref:kennedy:2016} defines stronger \index{synapse}synapses as ones that contribute more depolarisation to the neural membrane upon activation than weaker ones. Stronger \index{synapse}synapses have a higher probability of generating an action potential in the post-synaptic neuron. During activation, the pre-synaptic neuron release neurotransmitters that bind to the post-synaptic neuron~\cite{ref:khanacademy:synapse}. The frequent release of these molecules cause the \index{synapse}synapse to grow. Connections that grow over time yield stronger signals (learning), while connections that are weak propagate low intensity signals and vanish over time (forgetting). The ability of synapses to strengthen and weaken over time is known as \index{synaptic plasticity}\textit{synaptic plasticity}~\cite{ref:huttenlocher:1997}.

@@

The \acs{AN} implements a nonlinear mapping from $\mathbb{R}^{I}$ to
$\mathbb{R}^{T}$, usually in the ranges $[0,1]$ or $[-1,1]$, depending on the
activation function used~\cite{ref:engelbrecht:2007} and is given
in Equation~\eqref{eq:an_function_mapping} below.

@@

Input signals are $I$-dimensional vectors of numerical values that are obtained either through some environment sensor or from other \acp{AN}. Input signals are often referred to as \index{features}\textit{features} or \index{independent variables}\textit{independent variables}~\cite{ref:francis:2001}. Throughout this dissertation, a single input vector is referred to as a \textit{pattern}. Input data must first be pre-processed before it is presented to the \acs{AN}. Input pre-processing techniques are presented in the following sections.

@@

Input pre-processing improves the training process and contributes to the success of the practical application of the \acs{ANN}~\cite{ref:kuzniar:2017}. The primary purpose of input pre-processing is to modify the input data so that it can better match predicted output data. There are many techniques to consider when pre-processing input data. These techniques include methods to encode data to certain formats, to scale to the correct ranges and to deal with incomplete, invalid or irrelevant data. For the purposes of this dissertation, \textit{data encoding} and \textit{normalisation} techniques are considered.

@@

For qualitative data, the class labels have to be converted from textual to numerical representations. In general, class labels are encoded as numerical vectors~\cite{ref:srinidhi:2018, ref:brownlee:2017:one-hot}. One such encoding technique is referred to as \index{one-hot encoding}\textit{one-hot encoding}. Harris and Harris~\cite{ref:harris:2010} describe \index{one-hot encoding}one-hot encoding as a group of vector elements where the classes are represented with a single high element (represented by $1$) and all the others low (represented by $0$). Each element $y_k$, where $k$ is the $k$-th element of the encoded vector $\vec{y}$, uniquely refers to a class. An activation, represented by a value of $1$ at a given index thus represents a class associated with that index.

@@

Numerical data must be normalised and scaled appropriately for the \acs{AN}. In general, input data is normalised to a range that is appropriate for the \index{activation function}activation function being used by the \acs{AN}. Normalisation and scaling yields input data that is comparable to the output produced by the \acs{AN}. For the purposes of this disseration, the \index{min-max scaler}min-max scaler~\cite{ref:al:2006} and the \index{standard score scaler}standard score scaler~\cite{ref:jain:2005} are considered.

@@

The connection strength that synapses in the \acs{BN} have is modeled in the \acs{AN} as weight vectors of numerical values associated with each dimension of the input. Weights can either dampen or strengthen the input by some negative or positive numerical value. Changes in the weight associated with a feature changes the influence that that particular feature has on the predicted output. Finding the correct value for each weight, such that the \acs{AN} yields optimal output, is an optimisation problem~\cite{ref:thierens:1993}. Research has shown that weight initialisation plays an important role in the efficient training of \acp{ANN}~\cite{ref:thimm:1995}.

@@

Weight initialisation is the process by which candidate solutions (represented by the weights of the \acs{AN}) to the problem are ``placed'' in the search-space. Weight initialisation influences the speed of convergence, the probability of convergence and the generalisation capabilities of \acp{ANN}~\cite{ref:fernandez:2001}. Finding the optimal initialisation values for weights is non-trivial and can be seen as another optimisation problem~\cite{ref:de:2016, ref:erdogmus:2003, ref:yam:2000}.

Weight initialisation is dependent on the \index{activation function}activation function used. If weights are initialised as values that are too small, the vanishing gradients problem can occur~\cite{ref:hanin:2018}. If weights are initialised as values that are too big, output of the activation function would not be in the active range. This leads to unit saturation, and exploding gradients can occur~\cite{ref:hanin:2018, ref:yadav:2018}.

Many different weight initialisation techniques have been developed~\cite{ref:erdogmus:2003}. Some techniques have been developed to address the problems that are mentioned above~\cite{ref:yadav:2018}. For the purposes of this dissertation, focus is put on \index{random uniform sampling}random uniform sampling, \index{Glorot uniform sampling}\textit{Glorot uniform (Xavier)} sampling and \index{Glorot normal sampling}\textit{Glorot normal} sampling~\cite{ref:glorot:2010}. Brief discussions on each of these weight initialisation techniques are presented as follows.

@@

\index{random uniform sampling}Random uniform sampling initialises weights uniformly in the range $[\omega_{min}, \omega_{max}]$, written as $w_{i} \sim \textit{U} (\omega_{min}, \omega_{max})$, where the $\omega_{min}$ and $\omega_{max}$ are respectfully the lower and upper bounds of the uniform distribution. Suggested parameter values are $(-1, 1)$ or $(-0.5, 0.5)$~\cite{ref:nguyen:1990}.

@@

\index{Glorot normal sampling}Glorot normal sampling initialises weights by sampling from a truncated normal distribution centred on a mean of $0$ and with $\sigma = \sqrt{\frac{2}{I + K}}$, where $\sigma$ is the standard deviation of the distribution. $I$ and $K$ are respectfully the number of input and output units in the weight vector. \index{Glorot normal sampling}Glorot normal sampling has been shown to decrease training time~\cite{ref:glorot:2010}.

@@

A common way by which the \index{net input signal}net input signal is calculated is by means of \acp{SU}~\cite{ref:engelbrecht:2007}, which compute the \index{net input signal}net input signal as the weighted sum of all input signals as shown in Equation~\eqref{eq:summation_units} below.

@@

A bias/threshold term $\theta$ is introduced to help translate the output of the \index{activation function}activation function~\cite{ref:benitez:1997}. The value of $\theta$ can be learned during the training process along with the weights of the \acs{ANN}. In order to simplify equations, the input and weight vectors are augmented such that the input and hidden layers have an additional neuron/unit, called the \textit{bias unit}~\cite{ref:engelbrecht:2007}. The net input signal can then be rewritten to consider the bias unit, leading to an augmented net input signal.

@@

An \index{activation function}activation function is used to model the action potential of the \acs{AN}~\cite{ref:ziv:1994, ref:hodgkin:1952}. The \index{activation function}activation function takes as a parameter, the augmented \index{net input signal}net input signal. When the output produced by the activation function surpasses some threshold value $\tau$, we consider that neuron to have ``fired'' an output signal. \index{activation function}Activation functions thus model \textit{phase shift}. In the context of classification problems, \index{activation function}activation functions form decision boundaries between classes. In the context of regression problems, \index{activation function}activation functions try to approximate some function that maps the input data to some target data.

@@

\noindent
Many different \index{activation function}activation functions have been developed~\cite{ref:karlik:2011}. For the purposes of this dissertation, focus is put on the \acf{ReLU}~\cite{ref:jarrett:2009, ref:nair:2010}, the \acf{LReLU}~\cite{ref:maas:2013}, the \index{sigmoid}sigmoid~\cite{ref:lecun:1988} and the \index{hyperbolic tangent}hyperbolic tangent~\cite{ref:lin:2008} \index{activation function}activation functions.

@@

\noindent
An advantage of \acs{ReLU} is that it is not susceptible to the \index{vanishing gradients problem}vanishing gradients problem~\cite{ref:xu:2015, ref:maksutov:2018} which occurs when gradients in the first layers of a multi-layer \acs{ANN} approach zero, and have no effect on the training process.

@@

The \acf{LReLU} \index{activation function}activation function is a variant of the \acs{ReLU} \index{activation function}activation function that avoids zero gradients in the negative part of its argument by introducing a scaling parameter, $\alpha > 0$~\cite{ref:xu:2015}. Similar to \acs{ReLU}, \acs{LReLU} is not susceptible to the \index{vanishing gradients problem}vanishing gradients problem. However, \acs{LReLU} does not suffer from the \index{dying ReLU problem}dying \acs{ReLU} problem~\cite{ref:agarap:2018}, which occurs when the neurons that use \acs{ReLU} \index{activation function}activation functions become \textit{inactive} and only output $0$ due to a negative net input signal. The \acs{LReLU} \index{activation function}activation function is defined in Equation~\eqref{eq:leaky_relu} below.

@@

\noindent
By introducing a parameter $\alpha > 0$, negative \index{net input signal}net input signals will still yield non-zero activations, resulting in non-zero gradients and avoiding gradient saturation. Non-zero gradients are required by some heuristics such as \acs{SGD} in order to be able to effectively train \acp{ANN}~\cite{ref:hanin:2018}. An illustration of \acs{LReLU} with various values for $\alpha$ and $\theta = 0$ is given in Figure~\ref{fig:anns:activation_functions:leaky_relu}.

@@

The \index{sigmoid}sigmoid activation is the continuous differentiable approximation of the step function, which was used in the original perceptron model developed by~\citeauthor{ref:rosenblatt:1957}~\cite{ref:rosenblatt:1957}, and yields output in the range $(0, 1)$. The \index{sigmoid}sigmoid \index{activation function}activation function is defined in Equation~\eqref{eq:sigmoid} below.

@@

The \index{softmax}softmax function, also known as the \index{softargmax}\textit{softargmax}~\cite[p.~184]{ref:goodfellow:2016} or \index{normalised exponential function}\textit{normalised exponential function}~\cite{ref:bishop:2006} is a generalisation of the logistic function that converts a $K$-dimensional output vector $\vec{y}$ into a $K$-dimensional output vector $\vec{y^{'}}$ where each element $y^{'}_k$ is in the range $(0,1)$ and all elements sum up to $1$ as is shown in Equation~\eqref{eq:sum_after_softmax} below.

@@

\acp{ANN} are inspired by the biological brain. Our brains model biological neural networks with neuron counts in the hundreds of billions and synapse counts in the trillions. To emulate the computational capability of the human brain is not yet possible on modern day hardware.~\citeauthor{ref:sandberg:2008}~\cite{ref:sandberg:2008} approximates a computational requirement of 256 000 terabytes/s to emulate the entire brain. Despite our shortage in hardware capabilities, \acp{ANN} have been successfully applied to a range of problem classes.~\citeauthor{ref:engelbrecht:2007}~\cite{ref:engelbrecht:2007} summarises some common problems that are solved using \acp{ANN} and can be listed as follows.

\begin{itemize}
      \item \textbf{Classification}: Predicting the class of an input vector~\cite{ref:khan:2001}.

      \item \textbf{Pattern Matching}: Producing closely associated patterns based on an input vector~\cite{ref:cannady:1998, ref:kumar:1994}.

      \item \textbf{Pattern Completion}: Completing the missing parts of an input vector~\cite{ref:dayhoff:2001}.

      \item \textbf{Optimisation}: Producing optimal values of parameters in a optimisation problems~\cite{ref:specht:1991}.

      \item \textbf{Data Mining}: Feature discovery in large datasets~\cite{ref:singh:2009}.
\end{itemize}

@@

The input layer contains the input data to the \acs{ANN}. Since the input layer simply provides the input data, some literature do not consider the input layer as an actual part of the \acs{ANN}~\cite{ref:engelbrecht:2007}.

@@

The hidden layer contains a collection of ``hidden'' \acp{AN}, also referred to as \textit{hidden units} or \textit{nodes}. Hidden units are used if the target data is not linearly separable~\cite{ref:engelbrecht:2007}. It has been shown that \acp{ANN} that incorporate monotonically increasing differentiable \index{activation function}activation functions can approximate any continuous function with just one hidden layer, given that the hidden layer has enough hidden units~\cite{ref:hornik:1989}.

@@

The topology of the \acs{ANN} refers to the way that layers of \acp{AN} are connected to each other. There are many different topologies~\cite{ref:miikkulainen:2010}. For the purposes of this dissertation focus is put on a \index{fully connected topology}\textit{fully connected} topology where each \acs{AN} in one layer is connected to all \acp{AN} in the next, without any cycles~\cite{ref:zell:1994}.

@@

\acp{FFNN} were the first and simplest type of \acp{ANN} developed~\cite{ref:schmidhuber:2015} and implement input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement \index{fully connected topology}fully connected topologies. In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes. Depending on the optimisation algorithm used, error correction information can be ``backpropagated'' through the network. An illustration of a \acs{FFNN} is given in Figure~\ref{fig:ffnn} below.

@@

\index{supervised learning}Supervised learning is the process of training where the data that is presented to the \acs{FFNN} during training, includes the desired solution~\cite{ref:geron:2017}.  The \acs{FFNN} learns the mapping function from the input to the target output~\cite{ref:brownlee:2016}. The desired solutions are referred to as \textit{labels}. \index{supervised learning}Supervised learning can be used for both classification and regression problems.

The training data that is used during supervised learning, is split proportionally into a \textit{training} and \textit{validation} set. Data in the training set is used to train the \acs{FFNN}~\cite{ref:james:2013} and update the weights, while the validation dataset is used for hyper-parameter tuning, where the parameters of the training process are altered, but not the weights of the \acs{FFNN}.

@@

\index{overfitting}Overfitting describes a scenario where the trained \acs{FFNN} performs well on training data, but does not generalise well to never before seen data from the test set~\cite{ref:tetko:1995, ref:geron:2017}. Geron~\cite{ref:geron:2017} describes \index{overfitting}overfitting as the case where \acs{FFNN} is too complex relative to the noisiness of the training data.

@@

Underfitting describes a scenario where the \acs{FFNN} is not able to effectively learn the underlying structure of the training data~\cite{ref:tetko:1995, ref:geron:2017}. Geron~\cite{ref:geron:2017} describe underfitting as the case where the \acs{FFNN} is too simple relative to the underlying structure of the training data.\\
\\
\noindent
There are two types of supervised learning algorithms based on when weights are updated~\cite{ref:engelbrecht:2007}. These include stochastic training and batch
training.

@@

\index{stochastic training}Stochastic training, also known as \index{online learning}\textit{online learning}, is a \index{supervised learning}supervised learning variation whereby weights are adjusted after each training pattern presented. \index{stochastic training}Stochastic training benefits from shuffling data in the training dataset before presenting it to the \acs{FFNN}, in order to avoid overfitting or memorising the order in which patterns are presented~\cite{ref:engelbrecht:2007}. It has been shown that shuffling the training data can speed up convergence
\cite{ref:bengio:2012}.

@@

Research suggests a trade-off between stochastic and batch training by making use of mini-batches~\cite{ref:bengio:2012}. Mini-batch training is similar to batch training, however weights are updated after $\beta$ patterns have been presented, where $\beta$ is the mini-batch size.\\
\\
\noindent
Performance metrics are used during training on the training set and evaluation on the test set. There are many different performance measurements that can be used, however this dissertation focuses on performance measures related to loss. Loss is calculated using an error function. The following section presents the reader with more detail on the error functions that can be used during the training process.
